---
title: "Homework5"
author: "Weixi Chen"
date: "2/23/2022"
output: github_document
---

# homework 5

```{r}
# read the dataset we need
library('MASS') ## for 'mcycle'
library('manipulate') ## for 'manipulate'

y <- mcycle$accel
x <- matrix(mcycle$times, length(mcycle$times), 1)

data(mcycle)

plot(x, y, xlab="Time (ms)", ylab="Acceleration (g)")
```

## Question1: randomly split the mcycle data

```{r}
set.seed(1)
training_index <- sample(1:nrow(mcycle), dim(mcycle)[1]*0.75)
training_data <- mcycle[training_index, ]
validation_data <- mcycle[-training_index, ]
```


## Question2: Use Nadaraya-Waston method with the k-NN kernel function to predict the mean acceleration as a function of time

```{r}
## k-NN kernel function
## x  - n x p matrix of training inputs
## x0 - 1 x p input where to make prediction
## k  - number of nearest neighbors
kernel_k_nearest_neighbors <- function(x, x0, k=1) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
```

```{r}
## Make predictions using the NW method
## y  - n x 1 vector of training outputs
## x  - n x p matrix of training inputs
## x0 - m x p matrix where to make predictions
## kern  - kernel function to use
## ... - arguments to pass to kernel function
nadaraya_watson <- function(y, x, x0, kern, ...) {
  k <- t(apply(x0, 1, function(x0_) {
    k_ <- kern(x, x0_, ...)
    k_/sum(k_)
  }))
  yhat <- drop(k %*% y)
  return(yhat)
}
```

```{r}
# we can get the predictions by using the function
nadaraya_watson(as.matrix(training_data$accel), as.matrix(training_data$times), as.matrix(validation_data$times), kernel_k_nearest_neighbors)

# also we can change the tuning parameter k in 'kernel_k_nearest_neighbors' to get a series of prediction models.
```


## Question3: compute and plot the training error, AIC, BIC, and validation error

```{r}
# the training error

## loss function
## y    - train/test y
## yhat - predictions at train/test x
loss_squared_error <- function(y, yhat)
  (y - yhat)^2

k1 <- seq(1, 15, 1)

y_hat <- matrix(NA, nrow = dim(training_data)[1], ncol = length(k1))

for (k2 in k1){
  kernel_k_nearest_neighbors <- function(x, x0, k = k2) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
  
  y_hat[,k2] <- nadaraya_watson(as.matrix(training_data$accel), as.matrix(training_data$times), as.matrix(training_data$times), kernel_k_nearest_neighbors)
}

## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

training_error <- c()
for(x in 1:dim(y_hat)[2]){
  training_error <- c(training_error, error(training_data$accel, y_hat[,x]))
}

plot(k1, training_error, type = "b", xlab = "tuning parameter(k)", ylab = "training error")
```

```{r}
# AIC
k1 <- seq(1, 15, 1)

y_hat <- matrix(NA, nrow = dim(training_data)[1], ncol = length(k1))

for (k2 in k1){
  kernel_k_nearest_neighbors <- function(x, x0, k = k2) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
  
  y_hat[,k2] <- nadaraya_watson(as.matrix(training_data$accel), as.matrix(training_data$times), as.matrix(training_data$times), kernel_k_nearest_neighbors)
}

## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
aic <- function(y, yhat, d)
  error(y, yhat) + 2/length(y)*d

aic1 <- c()
for(x in 1:dim(y_hat)[2]){
  aic1 <- c(aic1, aic(training_data$accel, y_hat[,x], d = 1))
}

plot(k1, aic1, type = "b", xlab = "tuning parameter(k)", ylab = "AIC")
```

```{r}
# BIC
k1 <- seq(1, 15, 1)

y_hat <- matrix(NA, nrow = dim(training_data)[1], ncol = length(k1))

for (k2 in k1){
  kernel_k_nearest_neighbors <- function(x, x0, k = k2) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
  
  y_hat[,k2] <- nadaraya_watson(as.matrix(training_data$accel), as.matrix(training_data$times), as.matrix(training_data$times), kernel_k_nearest_neighbors)
}

## y    - training y
## yhat - predictions at training x
## d    - effective degrees of freedom
bic <- function(y, yhat, d)
  error(y, yhat) + log(length(y))/length(y)*d

bic1 <- c()
for(x in 1:dim(y_hat)[2]){
  bic1 <- c(bic1, bic(training_data$accel, y_hat[,x], d = 1))
}

plot(k1, bic1, type = "b", xlab = "tuning parameter(k)", ylab = "BIC")
```

```{r}
# Validation Error
k1 <- seq(1, 15, 1)

y_hat <- matrix(NA, nrow = dim(validation_data)[1], ncol = length(k1))

for (k2 in k1){
  kernel_k_nearest_neighbors <- function(x, x0, k = k2) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
}
  
  y_hat[,k2] <- nadaraya_watson(as.matrix(training_data$accel), as.matrix(training_data$times), as.matrix(validation_data$times), kernel_k_nearest_neighbors)
}

## test/train error
## y    - train/test y
## yhat - predictions at train/test x
## loss - loss function
error <- function(y, yhat, loss=loss_squared_error)
  mean(loss(y, yhat))

validation_error <- c()
for(x in 1:dim(y_hat)[2]){
  validation_error <- c(validation_error, error(validation_data$accel, y_hat[,x]))
}

plot(k1, validation_error, type = "b", xlab = "tuning parameter(k)", ylab = "Validation Error")
```


## Question4: 5-fold cross-validation

```{r}
library(plyr)
CV <- function(k, size, seed){
  cvlist <- list()
  set.seed(seed)
  n <- rep(1:k, ceiling(size/k))[1:size]
  temp <- sample(n, size)
  x <- 1:k
  dataseq <- 1:size
  cvlist <- lapply(x, function(x) dataseq[temp == x])
  return(cvlist)
}

cv_index <- CV(5, size = dim(mcycle)[1], seed = 1)
cv_index
```

```{r}
# test Error
k1 <- seq(1, 15, 1)

test_error <- matrix(NA, nrow = 5, ncol = length(k1))

for(index in 1:5){
for (k2 in k1){
  kernel_k_nearest_neighbors <- function(x, x0, k = k2) {
  ## compute distance betwen each x and x0
  z <- t(t(x) - x0)
  d <- sqrt(rowSums(z*z))

  ## initialize kernel weights to zero
  w <- rep(0, length(d))
  
  ## set weight to 1 for k nearest neighbors
  w[order(d)[1:k]] <- 1
  
  return(w)
  }
  
  training_data <- mcycle[-cv_index[[index]],]
  test_data <- mcycle[cv_index[[index]],]
  
  y_hat1 <- nadaraya_watson(as.matrix(training_data$accel), as.matrix(training_data$times), 
                                as.matrix(test_data$times), kernel_k_nearest_neighbors)
  test_error[index, k2] <- error(mcycle[cv_index[[index]],]$accel, y_hat1)
  
}
}

# The test error stores the test error for each piece of test data.
test_error
```


## Question5: plot the test error

```{r}
stand_deviation <- apply(test_error, 2, sd)
test_error1 <- apply(test_error, 2, mean)
plot(k1, test_error1, type = "b", xlab = "tuning parameter(k)", ylab = "test error")
segments(x0 = k1, y0 = test_error1 - stand_deviation/2, x1 = k1, y1 = test_error1 + stand_deviation/2)
```

## Question6: Interpret the figure

From the figure above, we can find out that the 5-fold cross validation test error lower down first and then it becomes higher again. Thus, we can use k = 9 as the tuning parameter in our model.