---
title: "Homework6"
author: "Weixi Chen"
date: "3/26/2022"
output: github_document
---

```{r}
library('randomForest')  ## fit random forest
library('dplyr')    ## data manipulation
library('magrittr') ## for '%<>%' operator
library('gpairs')   ## pairs plot
library('viridis')  ## viridis color palette
library('caret') ## 'createFolds'
```

```{r}
# load the dataset we need
vowel <- 
  read.table(url(
    'https://hastie.su.domains/ElemStatLearn/datasets/vowel.train'), sep = ",", header = TRUE)[,-1]
```

## Question1: Convert the type of response variable

```{r}
vowel %<>%
  mutate_at(c('y'), as.factor)  
```

## Question2: Review "randomForest" document

```{r}
?randomForest
```

## Question3: fit the random forest model

```{r}
fit <- randomForest(x = vowel[,-1], y = vowel$y,  
                    ntree = 500, mtry = 3, proximity=TRUE)

print(fit)          ## summary of fit object
plot(fit)           ## plot OOB MSE as function of # of trees
importance(fit)     ## variable importance 
varImpPlot(fit)     ## variable importance plot
```

## Question4: 5-fold CV and model tuning

```{r}
#split dataset
vowel_flds  <- createFolds(vowel$y, k=5)
print(vowel_flds)
sapply(vowel_flds, length)  ## not all the same length
```

### the number of variables randomly sampled as candidates at each split

```{r}
set.seed(2)
split_num <- c(3, 4, 5)
cverr <- rep(NA, length(split_num))
for (i in 1:length(split_num)) {
  cverr1 <- rep(NA, length(vowel_flds))
  for(j in 1:length(vowel_flds)) { ## for each fold
    
    ## get training and testing data
    vowel_train <- vowel[-vowel_flds[[j]],]
    vowel_test <- vowel[ vowel_flds[[j]],]
    
    ## fit random forest model to training data
    randomforest_fit <- randomForest(x = vowel_train[,-1], y = vowel_train$y,  
                    ntree = 500, mtry = split_num[i], proximity=TRUE)
    
    ## compute test error on testing data
    pre_tst <- predict(randomforest_fit, vowel_test)
    cverr1[j] <- sum(pre_tst != vowel_test$y)/length(pre_tst)
  }
  cverr[i] <- mean(cverr1)
}
cverr
```

So we should choose 3 as the number of variables randomly sampled as candidates at each split.

### the minimum size of terminal nodes

```{r}
set.seed(3)
node_size <- c(1, 5, 10, 20, 40, 80)
cverr <- rep(NA, length(node_size))
for (i in 1:length(node_size)) {
  cverr1 <- rep(NA, length(vowel_flds))
  for(j in 1:length(vowel_flds)) { ## for each fold
    
    ## get training and testing data
    vowel_train <- vowel[-vowel_flds[[j]],]
    vowel_test <- vowel[ vowel_flds[[j]],]
    
    ## fit random forest model to training data
    randomforest_fit <- randomForest(x = vowel_train[,-1], y = vowel_train$y,  
                    ntree = 500, mtry = 3, nodesize = node_size[i], proximity=TRUE)
    
    ## compute test error on testing data
    pre_tst <- predict(randomforest_fit, vowel_test[, -1])
    cverr1[j] <- sum(pre_tst != vowel_test$y)/length(pre_tst)
  }
  cverr[i] <- mean(cverr1)
}
cverr
```

So we should choose 1 as the minimum size of terminal nodes。

## Question5: compute misclassification rate

```{r}
# load the dataset we need
vowel_test <- 
  read.table(url(
    'https://hastie.su.domains/ElemStatLearn/datasets/vowel.test'), sep = ",", header = TRUE)[,-1]

vowel_test %<>%
  mutate_at(c('y'), as.factor)  
```

```{r}
set.seed(4)
randomforest_fit <- randomForest(x = vowel[,-1], y = vowel$y, 
                    ntree = 500, mtry = 3, nodesize = 1,  importance=TRUE, proximity=TRUE)
    
## compute test error on testing data
pre_tst <- predict(randomforest_fit, vowel_test)
mis_rate = sum(pre_tst != vowel_test$y)/length(pre_tst)
mis_rate
```

So the misclassification rate is 0.4004329.